Bandit Multiclass Classification

### 2019-01-09 ICML-submission/figures/pizza-slice.tex
Start ICML submission

  \coordinate (C) at (120:3);
  \coordinate [label={[xshift=-6mm, yshift=-4mm]$\gamma$}] (C1) at ($(120:3) + (0.75, 0.43301270189221932338)$);
  \coordinate [label={[xshift=+3mm, yshift=+2mm]$\gamma$}] (C2) at ($(120:3) - (0.75, 0.43301270189221932338)$);
  \draw[<->, >=latex, ultra thick] (C1) -- (C);
  \draw[<->, >=latex, ultra thick] (C2) -- (C);

  \coordinate (D) at (240:3);
  \coordinate [label={[xshift=+4mm, yshift=-8mm]$\gamma$}] (D1) at ($(240:3) + (-0.75, 0.43301270189221932338)$);
  \coordinate [label={[xshift=-5mm, yshift=-3mm]$\gamma$}] (D2) at ($(240:3) - (-0.75, 0.43301270189221932338)$);
  \draw[<->, >=latex, ultra thick] (D1) -- (D);
  \draw[<->, >=latex, ultra thick] (D2) -- (D);

>>>>

  \coordinate (C) at (120:3);
  \coordinate (C1) at ($(120:3) + (0.75, 0.43301270189221932338)$);
  \coordinate (C2) at ($(120:3) - (0.75, 0.43301270189221932338)$);
  \draw[<->, >=latex, ultra thick] (C1) -- (C);
  \draw[<->, >=latex, ultra thick] (C2) -- (C);
  \draw ($0.5*(C1) + 0.5*(C)$) node[above, xshift=-1mm] {$\gamma$};
  \draw ($0.5*(C2) + 0.5*(C)$) node[above, xshift=-1mm] {$\gamma$};

  \coordinate (D) at (240:3);
  \coordinate (D1) at ($(240:3) + (-0.75, 0.43301270189221932338)$);
  \coordinate (D2) at ($(240:3) - (-0.75, 0.43301270189221932338)$);
  \draw[<->, >=latex, ultra thick] (D1) -- (D);
  \draw[<->, >=latex, ultra thick] (D2) -- (D);
  \draw ($0.5*(D1) + 0.5*(D)$) node[below, xshift=-1mm] {$\gamma$};
  \draw ($0.5*(D2) + 0.5*(D)$) node[below, xshift=-1mm] {$\gamma$};

### 2019-01-09 ICML-submission/introduction.tex
Start ICML submission

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\cite{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a game between the learner and the adversary: at each timestep $t$,
the adversary chooses an example $(x_t, y_t)$ where the feature vector $x_t \in
\R^d$ is revealed to the learner and the correct label $y_t$ is kept secret by
the adversary. Upon receiving the feature vector $x_t$, the learner is asked to
make a prediction $\widehat{y}_t$. Immediately after making its prediction, the
learner receives a feedback. In contrast with the standard full information
setting (where the feedback given is the correct label $y_t$), here the feedback
is only a binary indicator of whether the prediction was correct or not. The
protocol of the problem is formally stated below.

@@

In this paper, we focus on the special case when the examples lie in $\R^d$ and
are \emph{linearly separable with a margin}.
\chicheng{I wonder if we should introduce the two notions of linear separability here,
because strong linear separability is also worth mentioning in the full-information setting.
}
\david{The formal definitions are in a section of their own.}

@@

The bandit feedback setting, however, is much more challenging. To understand
the differences, we dissect the notion of linear separability in the multiclass
setting. Beside the standard notion of linear separability, we define also
\emph{strong linear separability}. For binary classification, these two
definitions are identical. However, they differ for multiclass classification.
For multiclass classification with $K$ classes the standard notion of linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane. We state the definitions precisely in
Section~\ref{section:notions-of-linear-separability}.

For the case when the examples are strongly linearly separable, we design a
simple and efficient algorithm with expected number of mistakes at most $(K-1)
\lfloor (R/\gamma)^2 \rfloor$. The algorithm can be viewed as running $K$ copies
of the \textsc{Binary Perceptron} algorithm, one copy for each class.
Intuitively speaking, the factor $K-1$ is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We also
prove that any (possibly randomized) algorithm makes $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.

For the standard notion of linear separability, we propose several algorithms
with various time-complexity-versus-mistake-bound trade-offs. First class of
algorithms map the examples into a high-dimensional inner product space. In this
space the examples become \emph{strongly} linearly separable. This allows us to
run the algorithm for the strongly linearly separable case. All of these
algorithms can be kernelized. (Some of them have to be kernelized since the
underlying inner product space is infinite dimensional.) Kernelization does not
change the input-output behavior of the algorithm, it only affects its time
and memory complexity.

>>>>

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\citep{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a game between the learner and the adversary: at each timestep $t$,
the adversary chooses an example $(x_t, y_t)$ where the feature vector $x_t \in
\R^d$ is revealed to the learner and the correct label $y_t$ is kept secret by
the adversary. Upon receiving the feature vector $x_t$, the learner is asked to
make a prediction $\widehat{y}_t$. Immediately after making its prediction, the
learner receives a feedback. In contrast with the standard full information
setting (where the feedback given is the correct label $y_t$), here the feedback
is only a binary indicator of whether the prediction was correct or not. The
protocol of the problem is formally stated below.

@@

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and
are \emph{linearly separable with a margin}
(see Definition~\ref{def:linear-sep} for a formal description).
%\chicheng{I wonder if we should introduce the two notions of linear separability here,
%because strong linear separability is also worth mentioning in the full-information setting.
%}
%\david{The formal definitions are in a section of their own.}

@@

The bandit feedback setting, however, is much more challenging. To understand
the differences, we dissect the notion of linear separability in the multiclass
setting. Besides the standard notion of linear separability, we also consider
\emph{strong linear separability} (see Definition~\ref{def:strong-linear-sep} for a formal description). For binary classification, these two
definitions are identical.
However, they differ for multiclass classification with $K \geq 3$.
For multiclass classification with $K$ classes, the standard notion of linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane. We state the definitions precisely in
Section~\ref{section:notions-of-linear-separability}.

For the case when the examples are strongly linearly separable, we design a
simple and efficient algorithm with an expected number of mistakes at most $(K-1)
\lfloor (R/\gamma)^2 \rfloor$. The algorithm can be viewed as running $K$ copies
of the \textsc{Binary Perceptron} algorithm, one copy for each class.
Intuitively speaking, the factor $K-1$ is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We also
prove that any (possibly randomized) algorithm makes $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.

For the standard notion of linear separability, we propose several algorithms
with various time-complexity-versus-mistake-bound trade-offs.
\chicheng{I wonder if we should just present the results using the rational kernel,
 as results using other kernels are nonadaptive,
 and achieve qualitatively similar mistake bounds.}
First class of
algorithms map the examples into a high-dimensional inner product space. In this
space the examples become \emph{strongly} linearly separable. This allows us to
run the algorithm for the strongly linearly separable case. All of these
algorithms can be kernelized. (Some of them have to be kernelized since the
underlying inner product space is infinite dimensional.) Kernelization does not
change the input-output behavior of the algorithm, it only affects its time
and memory complexity.

### 2019-01-09 ICML-submission/main-supplementary.tex
Start ICML submission

\documentclass[12pt]{article}

% LaTeX packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[singlelinecheck=false]{caption}
\usepackage{fullpage}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{thmtools}
\usepackage{tikz}
\usepackage{color}
\usepackage{comment}
\usepackage{commath}
\usepackage{dsfont}

% TikZ packages
\usetikzlibrary{arrows}
\usetikzlibrary{angles}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{cd}  % commutative diagrams
\usetikzlibrary{intersections}
\usetikzlibrary{patterns}
\usetikzlibrary{shapes}
\usetikzlibrary{through}

% Math environments
\newtheorem{definition}{Definition}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{corollary}[definition]{Corollary}

% Math symbols and commands
\newcommand{\calF}{\mathcal{F}}
\newcommand{\one}{\mathds{1}}
\newcommand{\R}{\mathbb{R}}  % set of real numbers
\newcommand{\indicator}[1]{\mathbf{1}\left[#1 \right]} % indicator
\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle} % inner product
%\newcommand{\norm}[1]{\left\| #1 \right\|}  % norm of a vector or a matrix
%\DeclareMathOperator*{\deg}{deg}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\Exp}{\mathbf{E}}  % expected value
\DeclareMathOperator{\diff}{d \!} % differential
\DeclareMathOperator*{\polylog}{polylog}

\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkred}{rgb}{0.7,0,0}
\definecolor{teal}{rgb}{0.3,0.8,0.8}
\definecolor{orange}{rgb}{1.0,0.5,0.0}
\definecolor{purple}{rgb}{0.8,0.0,0.8}
\definecolor{blue}{rgb}{0.0,0.0,1.0}
\newcommand{\kibitz}[2]{{\textcolor{#1}{\textsf{\footnotesize #2}}}}
\newcommand{\alina}[1]{\kibitz{darkred}{[AB: #1]}}
\newcommand{\david}[1]{\kibitz{darkgreen}{[DP: #1]}}
\newcommand{\balazs}[1]{\kibitz{purple}{[BS: #1]}}
\newcommand{\deva}[1]{\kibitz{teal}{[DT: #1]}}
\newcommand{\chenyu}[1]{\kibitz{orange}{[CW: #1]}}
\newcommand{\chicheng}[1]{\kibitz{blue}{[CZ: #1]}}

\title{Online Multiclass Linear Classification with Bandit Feedback}

\author{
Alina Beygelzimer \and
D\'avid P\'al \and
Bal\'azs Sz\"or\'enyi \and
Devanathan Thiruvenkatachari \and
Chen-Yu Wei \and
Chicheng Zhang
}

@@

\maketitle

\begin{abstract}
We study the problem of efficient online multiclass linear classification with
bandit feedback, where all examples belong to one of $K$ classes and lie in the
$d$-dimensional Euclidean space. Previous works have left open the challenge of
designing efficient algorithms with finite mistake bounds when the data is
linearly separable by a margin $\gamma$. In this work, we take a first step
towards this problem. We consider two notions of linear separability:
strong linear separability and weak linear separability.

\begin{enumerate}
\item Under the strong linear separability condition, we design an efficient
algorithm that achieves a near-optimal mistake bound of $O\left(\frac{K}{\gamma^2} \right)$.

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a infinite-dimensional feature mapping that transforms weak
linear separable examples to strong linear separable examples, which in turn is inspired by
\cite{Klivans-Servedio-2008}'s work on learning intersection of halfspaces.
\end{enumerate}
\end{abstract}

\input{introduction}
\input{related-work}
\input{notions-of-linear-separability}
\input{algorithm-for-strongly-linearly-separable-data}
\input{from-linear-separability-to-strong-linear-separability}
\input{proofs-of-polynomial-approximation-theorems}
\input{nearest-neighbor-algorithm}
\input{np-hardness-of-the-weak-labeling-problem}
\input{mistake-lower-bound-for-ignorant-algorithms}

\bibliographystyle{plainnat}
\bibliography{biblio}

\appendix

\input{appendix-multiclass-perceptron}

>>>>

\documentclass{article}

\input{commands}

@@

\input{title-and-abstract}
\input{introduction}
\input{related-work}
\input{notions-of-linear-separability}

\bibliography{biblio}
\bibliographystyle{icml2019}

% For actual submission remove everything below.
% Use appendices only in supplementray material.
\appendix
\clearpage
